---
title: "Make Data Frames"
author: "Homer White"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 2
---

## Attach Packages

```{r include = FALSE}
library(tidyverse)
library(lubridate)
```

## Import

```{r}
racers <- readr::read_csv("racers.csv")
teams <- readr::read_csv("teams.csv")
results <- readr::read_csv("results.csv")
```

The data table `racers` contains the following variables:

* an ID for the runner, provided by the race organizer when the runner registered;
* the name of the runner (last, first);
* the runner's home address (note that some addresses have only four-digits for the zip-code);
* the sex of the runner (M = male, F = female);
* the runner's date of birth;
* an ID for the runner's team (note that some runners are "independent"; their team ID is `NA`).

The data table `teams` contains the following variables:

* the ID for the team;
* the name of the team.

The data table `results` contains the following variables:

* the ID for the runner;
* the runner's time to complete the marathon.


## Cleaning

Your first task is to clean the `racers` data.  Use `racers` to create a new data table named `Racers`:

```{r}
Racers <- 
  racers %>%
  tidyr::extract(name, into = c("last_name", "first_name"),
                 regex = "(\\w+), (\\w+)") %>%
  mutate(dob = as_date(mdy(dob))) %>%
  select(id, first_name, last_name, sex, dob, team)
```


Now clean the `results` table.  The primary goal here is to convert the given times, which are strings, into a form that permits them to be compared numerically.  You need to create a new data table called `Results` that consists of the variables in results along with a new variable `seconds` that gives the completion time in seconds.

```{r}
Results <-
  results %>%
  mutate(seconds = as.numeric(lubridate::hms(time)))
```


### Joining

Join `Racers` with `Results` and with `teams` to get a new data table called `Marathon`.

```{r}
Marathon <-
  Racers %>%
  inner_join(Results, by = "id") %>%
  left_join(teams, by = c("team" = "teamID"))
```

## Save

Now save the data in the right place:

```{r}
save(Racers, file = "../data/Racers.rda")
save(Results, file = "../data/Results.rda")
save(teams, file = "../data/teams.rda")
save(Marathon, file = "../data/Marathon.rda")
```











